---
title: Tuning the GBM learning rate
author: "Sidharth Gupta"
date: \today
output:
   html_document:
    css: css/code.css
    toc: true
    toc_depth: 6
    highlight: default
    citation_package:
    keep_tex: false
    keep_md: true
    fig_caption: true
    latex_engine: pdflatex
runtime: shiny
fontsize: 11pt
geometry: margin=1ine
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setup, tidy = TRUE, message = FALSE, warning = FALSE, error = FALSE, results='hide'}
library(dplyr)
library(magrittr)
library(gridExtra)
library(ggplot2)
library(h2o)
library(gridExtra)
library(futile.logger)
library(kableExtra)
source("h2o/source-for-rmd.R")
SEED=123456
h2o.init()
```

```{r, echo=FALSE}
# KNITR CONTROL OPTIONS:
LOAD.DATA=FALSE # Once you run the NB and loaded the data you may want to make this false
LOAD.DISK.MODELS=TRUE # Once you run the NB the models will be loaded to disk, you might want to kill this option then
```

## Tuning the learning rate
In this vignette I will try and get a feel for the learning rate by training many GBM models on just the learning rate. Once these models are trained I will look at the scoring history and what learning rate does well.

### Load the data
First lets use the wrapper routines and prepare some data.

```{r data, cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, eval=LOAD.DATA==TRUE, results='hide'}
# The function for reading the data does a few things by default. It makes categorical  variables where ever necessar, and log transforms the large variables like for taxes etc. See documentation of that function.
list[tr.baseline, pr.baseline, vtreatIdentityFn, tplan.NULL] =  prepareDataWrapper.h2o.gbm.baseline()

independent.vars = colnames(tr.baseline %>% select(-logerror))

list[XYTrain.h2o, XYTest.h2o] = h2o.splitFrame(as.h2o(tr.baseline), ratios = 0.85, destination_frames=c("XYTrain.h2o", "XYTest.h2o"), seed=SEED)

```

## Defining the tree parameters
There are two data sets one used in the training and the other held out as a validation set. I created a grid with defaults, especially a large number of trees. But becasue we have early stopping we should be ok.

```{r hyperparams_search_crit}
hyper_params_search_crit.learnrate.tune = function(
            ntrees_opts = 10000, # early stopping will stop the earlier.
            max_depth_opts = 10,  # Start with 6 for now
            min_rows_opts = 5,   # How many leaf nodes to average over, doing the log of # rows would be ok
            learn_rate_opts = 0.01,
            learn_rate_annealing = 1,
            sample_rate_opts = 0.9,
            col_sample_rate_opts = 0.9,
            col_sample_rate_per_tree_opts = 0.9,
            strategy="RandomDiscrete") {

    list[search_criteria, hyper_params] =  hyper_params_search_crit(
            ntrees_opts = ntrees_opts,
            max_depth_opts = max_depth_opts,
            min_rows_opts = min_rows_opts,
            learn_rate_opts = learn_rate_opts, # from the previous
            learn_rate_annealing = learn_rate_annealing,
            sample_rate_opts = sample_rate_opts,
            col_sample_rate_opts = col_sample_rate_opts,
            col_sample_rate_per_tree_opts = col_sample_rate_per_tree_opts,
            strategy=strategy)
    return(list(search_criteria, hyper_params))
}

```


## Fitting the model
Here is the wrapper written to do the model fit.
```{r}

baseline.fit = function(hp, sc, grid_id = "learning_rate_search", seed=123456, ...) {
        h2o.grid.helper(
                        h2o.getFrame("XYTrain.h2o"),
                        h2o.getFrame("XYTest.h2o"),
                        independentCols=independent.vars,
                        hyper_params = hp,
                        search_criteria = sc,
                        algorithm="gbm",
                        grid_id=grid_id,
                        seed = seed,
                        ...
                        )
}
target_learning_rates = seq(0.01, 0.05, 0.01)
list[search_criteria.gbm.learnrate, hyper_params.gbm.learnrate] =
    hyper_params_search_crit.learnrate.tune(learn_rate_opts=target_learning_rates)
```

Now lets train a grid with different learning rates using crossvalidation and using the holdout data set as the guiding star to evaluate the best learning rate.

```{r training, cache=TRUE, results="hide"}
# Randomize and search for the best learning rate.
fit = baseline.fit(
       hp=hyper_params.gbm.learnrate,
       sc=search_criteria.gbm.learnrate,
       grid_id = paste("learning_rate_search"),
       seed = SEED, 
       nfolds=5)

```
## Analysis
Lets see what are the top learning rates:
```{r}
kable(fit@summary_table %>% data.frame %>% select(learn_rate, residual_deviance) %>% arrange(residual_deviance))
```

### Plotting the individual learning rate.
Save the models, optionally for later use
```{r, eval=FALSE}
LOCAL.RESULTS.DIR=paste0(getwd(), "/results/learn_rate_2018_15_04_gbm")
if(!dir.exists(LOCAL.RESULTS.DIR)) 
    dir.create(LOCAL.RESULTS.DIR)
h2o.gridSaver(c("learning_rate_search"), results.dir=LOCAL.RESULTS.DIR)
```

Optionally load the grid models for evaluation
```{r, eval=LOAD.DISK.MODELS}
LOCAL.RESULTS.DIR=paste0(getwd(), "/results/learn_rate_2018_15_04_gbm")
#models = h2o.gridLoader(Map(function(x) paste0("learning_rate_search_",x) , 1:NGrids), results.dir=LOCAL.RESULTS.DIR)
models = h2o.gridLoader(c("learning_rate_search"), results.dir=LOCAL.RESULTS.DIR)[, 1]

```

Or extract the individual models of each of the cross validated folds.
```{r modellist, cache=TRUE, eval=FALSE}
models = Map(h2o.getModel, fit@model_ids) 
```

```{r}
length(models)
```

Extract the score histories of each of the models into a data frame.
```{r, results="hide"}

getRMSEScoreHistory = function (models) {
    # Given a list of model ids get a data frame with its RMSE score history
    score.history = data.frame()
    # Learning rate of 0.03 seems to be ok.
    for (i in 1:length(models)) {
        model.rmse = h2o.rmse(h2o.performance(models[[i]]))
        lr = models[[i]]@parameters$learn_rate
        model.rmse.rate = h2o.scoreHistory(models[[i]]) %>%
            data.frame %>%
            pull(validation_rmse)
        score.history = rbind(score.history,
                              data.frame(
                                         ntrees = 1:length(model.rmse.rate),
                                         rmse = model.rmse.rate,
                                         id = i,
                                         lr = lr
                                         )
                              )
    }
    score.history
}

# For each model list in models extract the histories
score.history = getRMSEScoreHistory(models)
```

Plot the individual learning rate curves

```{r plots, fig.width=10, fig.height=10}
ggplot(score.history) +
        geom_line(aes(y=rmse, x=ntrees, color=as.factor(lr)),
                  position=position_dodge(1)) +
        theme(
          legend.position="bottom"
        )

```

## Conclusion and a note on learning rate variances

The learning rates begin to increase a bit at the end i.e. overshoot the optimum. The learning rates of 0.01-0.02 tended to perform better. I also repeated this experiment many times and evaluate the variance of the learning rate. 

```{r, cache=T, eval=FALSE, results-'hide'}
NGrids = 10
cls = registerCluster(max.cores.per.machine = 6)

fit.bootstrapped = foreach(x = 1:NGrids, .packages= c("h2o")) %dopar% {
              h2o.connect()
              baseline.fit(
                       hp=hyper_params.gbm.learnrate,
                       sc=search_criteria.gbm.learnrate,
                       grid_id = paste("learning_rate_search_", x, sep=''),
                       seed = x)
}

```

```{r, results='hide', echo=F, eval=F}

LOCAL.RESULTS.DIR=paste0(getwd(), "/results/learn_rate_2018_15_04_gbm.bootstrapped")
if(!dir.exists(LOCAL.RESULTS.DIR)) 
    dir.create(LOCAL.RESULTS.DIR)
h2o.gridSaver(Map(function(x) paste0("learning_rate_search_", x) , 1:NGrids), results.dir=LOCAL.RESULTS.DIR)
```

```{r, results='hide', echo=FALSE, eval=LOAD.DISK.MODELS==TRUE}
LOCAL.RESULTS.DIR=paste0(getwd(), "/results/learn_rate_2018_15_04_gbm.bootstrapped")
models = h2o.gridLoader(Map(function(x) paste0("learning_rate_search_",x) , 1:NGrids), results.dir=LOCAL.RESULTS.DIR)
```

 
Extract the combined score history and plot the grids
```{r plots.grid, echo=FALSE, fig.width=10, fig.height=10}

models.lst = sapply(X = 1:NGrids,
                    FUN=function(x) {
                        grid.id = h2o.getGrid(paste("learning_rate_search_", x, sep=''))
                        sapply(X=grid.id@model_ids, FUN=h2o.getModel)
                    })
score.history.lst = sapply(X = 1:NGrids,
                           function(x) getRMSEScoreHistory(models.lst[, x]))

score.history.combined = Reduce(function(x, init) {
               rbind(x, init)
          }, Map(function(i) {
                score.history.lst[, i] %>%
                    data.frame %>%
                    mutate(grid_id=i)
            }, 1:NGrids)
)

score.history.combined %>% group_by(grid_id, lr) %>% 
    dplyr::filter(ntrees==max(ntrees)) %>%
    data.frame %>% group_by(grid_id) %>% 
    dplyr::filter(rmse ==min(rmse)) %>%
    arrange(rmse) %>% kable

# I initially tried working with a lits of plots instead of plotting the data directly. Had trouble sizing the plots properly. So backed off and instead used a combined version of score history with vanialla ggplot.
min_y_lim = score.history.combined %>% pull(rmse) %>% min
max_y_lim = score.history.combined %>% pull(rmse) %>% max

ggplot(score.history.combined) +
    facet_wrap(~ grid_id, ncol=2) +
        geom_line(aes(y=rmse, x=ntrees, color=as.factor(lr)),
                  position=position_dodge(1)) +
        ylim(c(min_y_lim, max_y_lim)) +
        theme(
          legend.position="bottom",
          strip.background = element_blank(),
          strip.text.x = element_blank()
        )

```

These experiments show that in 6/10 cases rates of 0.1-0.3 do better than others. But there is still significant variance. Next lets look at annealing to improve on this.
